A Formal Verification Framework for Tezos Smart Contracts Based on Symbolic Execution

### 2024-06-04 papers/static-checker/aplas/paper.tex
The last version

%
%\authorrunning{Ha Doan, P. Thiemann}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
%\institute{University of Freiburg, Germany \\
%  \email{\{doanha,thiemann\}@informatik.uni-freiburg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
We developed SCV, a verification tool that is geared towards contracts implemented in Michelson, the smart contract language of the Tezos blockchain. SCV relies on symbolic execution to derive verification conditions that can be checked automatically against the user's specifications by employing an SMT solver. These specifications are written in a domain-specific language. We present the fundamental principles behind the design and development of SCV. By pruning states during symbolic execution, SCV addresses the problem of state explosion. This approach has enabled us to conduct two case studies on real-life contracts, demonstrating the expressiveness of our tool.
\keywords{Smart Contract, Blockchain, Formal Verification, Symbolic Execution}
\end{abstract}
\section{Introduction}
\label{sec:introduction}
Since smart contracts are permanently stored on the blockchain, once a smart contract is deployed, it cannot be changed. Regrettably, numerous instances have been documented where vulnerabilities in smart contracts were identified by the blockchain community, resulting in substantial financial losses~\cite{dao,wallethack}. The distinctive execution model of smart contract languages has given rise to unforeseen errors due to a lack of familiarity with their complexities. These observations, combined with the immutability of contracts, underscore the need to ensure their correctness prior to deployment.

We have developed SCV, a formal verification tool for smart contracts on the Tezos blockchain \cite{tezos-whitepaper}, written in Michelson \cite{michelson}. SCV relies on symbolic execution backed by an SMT solver to simulate the implementation of the smart contract language, helping to detect subtle errors. Additionally, SCV comes with a domain-specific language to precisely specify smart contract properties, tailored to Michelson's execution model and inspired by Z3. Each specification consists of Michelson code, a precondition, and a postcondition with the usual partial correctness semantics. Our symbolic interpreter calculates a postcondition predicate from the specified precondition. The SMT solver Z3 \cite{z3} checks that this predicate implies the specified postcondition.

%We have implemented approximately 85\% of the Michelson instructions, including most of the major-use instructions. The remaining instructions were not implemented primarily due to time constraints rather than technical limitations.

%One of the main challenges with formal verification is that users are often required to specify a system and express its properties in a specialized specification language. This requirement poses two problems: it demands expertise in formal verification from the specifier and introduces a gap between the actual code and its specification. In our framework, the actual Michelson code is used and users are not obliged to possess formal verification expertise. A basic understanding of Michelson data types and their usage when interacting with a smart contract on the Tezos blockchain is sufficient.

While there is already a considerable body of work applying symbolic execution to the verification of smart contracts \cite{securify, manticore, kevm, park} (see Section~\ref{sec:related-work} for details), most of these efforts are designed for Ethereum, where standard techniques for symbolic execution of imperative programs are applicable. Such groundwork does not exist for Michelson, which is a functional language running on a stack machine (see Section~\ref{sec:background}) with a sophisticated instruction set that includes high-level loops. For that reason, we had to develop symbolic execution for Michelson from the ground up.

@@

Since Michelson operates on a stack, we assume that the initial stack and the output stack each have only one element, as specified in the input term and the output term, respectively. The syntax allows us to specify both incomplete and complete smart contracts. For a complete smart contract, the output should be a pair consisting of an operation list and the storage's state after the completion of the code. For verification purposes, we elide the operation list and represent the output solely by the term of the value in the storage after execution concludes. %Michelson code, whether composed by users, obtained from the Tezos blockchain, or compiled from higher-level languages, can be seamlessly used within our framework.

@@

pre- and post-conditions, where the precondition states the assumption

@@

Let us have a look at the specification for the auction
contract that models an online auction in Listing \ref{lst:auction-contract-specification}. The contract has two entry points, \lstinline|%bid| and \lstinline|%close|, which serve as functions for bidding and closing the contract, respectively. The smart contract storage is a pair containing, as its first element, a boolean value \lstinline|open| that indicates whether the contract is still open for bidding or already closed. The second element is another nested pair, where the first element \lstinline|highest| stores the highest bidder's address, and the second element \lstinline|owner| is the contract owner's address. The current highest bid is reflected in the balance of the contract before the next bid is invoked. We omit the Michelson code of the smart contract due to its length.  For the \lstinline/%bid/ entrypoint, the parameter is \lstinline/Unit/ of type \lstinline/unit/.
The current storage is described by the \lstinline|storage| phrase.

@@

\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification},label={lst:auction-contract-specification},numbers=left]
mcontract auction = spec 
 storage := Pair (open: bool) 
                 (Pair (highest: address) (owner: address))
 entrypoint %bid
  code := {...}
  parameter := Unit 
  pre-condition := open = true && Amount > pre_Balance
   && get_contract (unit, highest) = Some (c: contract unit)
  post-condition := open = true && post_highest = Sender 
   && post_Balance = Amount && post_owner = owner 
   && transfer_token (Unit, pre_Balance, c);
 entrypoint %close
  ...
 (%create -> %bid) with ...
 | (%create  -> %close) with ...
 | (%bid -> %bid) with (Amount > pre_Balance) 
  && get_contract (unit, highest) = Some (c: contract unit)
 | (%bid -> %close) with (Sender = owner) 
  && get_contract (unit, owner) = Some (c: contract unit)
 | not (%close -> %bid) | not (%close -> %close)
\end{lstlisting}

%There are several properties that we want to guarantee for this contract. 
The first property is that if a user calls the \lstinline|%bid| entrypoint when the contract is still open for bidding and the bid is higher than the current bid (as specified in the \lstinline|pre-condition| field at line 7), and the address of the highest bidder can be cast to a real contract represented by the variable \lstinline|c| of type \lstinline|contract unit| on the blockchain (line 8), then the bid should succeed. In this case, the highest bidder must be updated to the sender, which is the address of the caller (line 9 in the \lstinline|post-condition| field), and the balance of the contract should be the amount sent along with the call (line 10). Moreover, the previous highest bidder should receive their bid back, which means the auction contract should transfer the \lstinline|pre_Balance| to the previous bidder, \lstinline|highest| (line 11). Another property guarantees that the auction owner's address should not be changed after calling the \lstinline/bid/ entrypoint (line 10). 
\subsection{Entrypoint relations}
\label{sec:entrypoint-relations}
Since the order of smart contract
calls can be crucial, we specify ordering constraints among calls with
 entrypoint relations. An entrypoint relation can be specified as
\lstinline/%a -> %b with p/, which starts with the name
\lstinline/%a/ of the entrypoint and is followed by the keyword
\lstinline/with/ in case conditions are specified as predicates. The
following symbol \lstinline/->/ indicates that from the entrypoint
that satisfies the condition \lstinline/p/, it is possible  to call
the entrypoint \lstinline/%b/. One specification can
contain several entrypoint relations concatenated by
\lstinline/|/. There is a special entrypoint named \lstinline/create/
that indicates the contract state after deployment.

Let us return to the auction contract. It is important to guarantee
that  the bid call can happen successfully after another if the new bid is higher than the previous
one (lines 16-17). The  \lstinline/%close/ entrypoint can be called
anytime after bidding if the caller is the contract owner (lines 18-19). However, it
should not be possible to call the  \lstinline/%bid/ entrypoint after
the auction is closed, and, of course, it cannot be closed
again if it is already closed (line 20).  Furthermore, contract deployment sets the  \lstinline/open/ flag to
\TRUE, indicating that the contract is open for bidding, i.e., users
should be able to bid on the contract or the contract owner should be
able to close it. Otherwise, the owner's original
deposit is lost. This sequence  illustrates
the life cycle of the auction contract, which is depicted in Figure
\ref{fig:auction-life-cycle}. Other properties related to the \lstinline|%close| entrypoint and the life cycle are omitted here, but the full specification can be found in the Appendix and in \cite{}.

@@

\section{Static Checker}
\label{sec:static-checker-we}
The architecture of the static checker is illustrated in Figure~\ref{fig:architecture-of-static-checker}. Users write specifications in the SCV DSL, which are then parsed into an Abstract Syntax Tree (AST) in OCaml. The input term along with the code is passed to the symbolic interpreter, while the output term and the pre- and post-conditions are fed to the static checker. The symbolic interpreter, which is fully typed and rejects ill-typed programs, then runs symbolic execution on the code with the provided input, where the initial stack is derived from the input term.

@@

Consider the specification of the \lstinline/sub/ contract, the results of
the symbolic execution are either a failwith state with the path
condition \lstinline/x/$<$\lstinline/y/,  or a state in which the
final stack has only one element, which is a symbolic variable
\lstinline/sbvar/ of type integer, with the path condition stating
that the variable \lstinline/x/ is greater than or equal to
\lstinline/y/ and  \lstinline/sbvar/ is equal to
the substract of \lstinline/x/ and \lstinline/y/.

@@

\subsection{Output check}
\label{sec:output-check}
For an incomplete smart contract, the output term of the symbolic execution has the form of a term \Term\ and its type \TY. For a complete smart contract, it takes the form (\PAIR\ \VOPERATIONLIST\ \VSTORAGE, \TPAIR\ (\TOPERATIONLIST) \TYS) \STACKCONCAT\ \EMPTYSTACK, where \VSTORAGE\ represents the storage and \VOPERATIONLIST\ represents the operation list. Given an output term specified as $\tau$, the checker scans all reachable final states that do not contain a failwith stack and checks whether the output term matches. In the former case, it matches $\tau$ with the term \Term. In the latter case, it matches with \VSTORAGE.

For instance, in the \lstinline|sub| contract, the output check ignores the states where the failwith occurs and verifies whether the output term \lstinline|z| matches with the symbolic output term in the other state. It succeeds for this example as expected. However, this test could fail if, for example, the output term is incorrectly specified as a variable of a type that does not match the actual result.

\subsection{Fail conditions}
\label{sec:fail-conditions}
A Michelson contract can fail due to conditions introduced by the programmer, triggered by the \FAILWITH\ instruction along with an error message. Failures play a crucial role in smart contract programming to ensure that certain conditions lead to contract termination, such as unauthorized caller attempts (access control properties) or invalid arguments. Various smart contract standards, including \cite{erc, fa}, mandate specific fail conditions.

@@

Any contract aiming to implement the FA 1.2 standard must incorporate the functions specified in Listing~\ref{lst:fa12-interface}. We verify USDtz \cite{usdtzimplemetation, tzstatsusdtz}, an implementation of the FA 1.2 standard. The goal is to ensure that the implementation aligns with the standard, verifying its correctness. USDtz is designed to maintain a stable value by pegging it to a fiat currency like USD or EUR, and it has been operational on the Tezos blockchain for a significant period.

@@

The \lstinline/transfer/ entrypoint credits the account associated with the address provided in the \lstinline/to/ parameter while debiting the account corresponding to the \lstinline/from/ parameter with the specified \lstinline/value/ amount. According to the FA 1.2 standard, this call must fail under two specific conditions: (1) The entrypoint must fail with the error \lstinline/NotEnoughBalance/ if the sender's account lacks sufficient funds. (2) If the sender's account does not have permission to withdraw the specified \lstinline/value/ amount, the call will fail with the \lstinline/NotEnoughAllowance/ error. To verify these requirements, we utilize the failwith condition feature of our tool and examine the returned results, some of which are outlined in Listing~\ref{lst:transfer-udstz-contract}. The fail condition results confirm these stipulations.
\begin{lstlisting}[float,captionpos=b,caption={The not enough balance fail condition for the \lstinline/transfer/ entrypoint},label={lst:transfer-udstz-contract},numbers=left]
not (paused = true) 
&& get_big_map (from, ledger) = Some sbvar 
&& sbvar = Pair sbvar1 sbvar2 
&& not (sbvar1 >= value) : 'NotEnoughBalance'
\end{lstlisting}
Another essential property is ensuring that the allowance of the sender's account is decremented and the receiver's account is incremented by the \lstinline/value/ that has been sent. This specification is verified to ensure compliance with the contract's intended behavior.

The \texttt{approve} entrypoint grants the spender the right to transfer a specified amount from the owner's account. The contract prohibits changing the allowance value from a non-zero value to another non-zero value to prevent a known attack vector \cite{attack-vector}. This attack exploits the ability to spend the entire allowance before an update and then again after, resulting in a larger amount being spent than allowed. Therefore, the contract call must fail if both the parameter \lstinline/value/ and the current allowance of \lstinline/spender/ are non-zero. We verify this critical property by generating fail conditions to check whenever such a non-zero value update results in a fail condition as Listing \ref{lst:approve-udstz-contract}.
\begin{lstlisting}[float=tp,captionpos=b,caption={The non-zero value fail condition for the \lstinline/approve/ entrypoint},label={lst:approve-udstz-contract},numbers=left]
not (paused = true) 
&& get_big_map(Sender, ledger) = Some sbvar 
&& sbvar = Pair sbvar1 sbvar2 
&& get_map(spender, sbvar2) = Some sbvar3 
&& not(value = 0) && not(sbvar3 = 0): 'UnsafeAllowanceChange'
\end{lstlisting}

When the update is allowed, the updated value must be correctly recorded in the storage. This can be divided into two sub-properties: (1) if the parameter \lstinline/value/ is zero, then the corresponding allowance value of the \lstinline/spender/ from the owner (which is \SENDER\ in this call) should be zero, and (2) if \lstinline/value/ is non-zero and the allowance of the \lstinline/spender/ is zero, then the \lstinline/value/ should be recorded in the storage after the call. We specify these properties using pre- and post-conditions, as illustrated in Listing~\ref{lst:specification-approve-1} for the first property.

@@

We specified and verified all the properties mentioned above and others. The results confirm that these properties are guaranteed, ensuring the correctness of the smart contract code.
\begin{lstlisting}[float=tp,captionpos=b,caption={Specification of the \lstinline/approve/ entrypoint (Property 1)},label={lst:specification-approve-1},numbers=left]
entrpoint %approve
 parameter := Pair (spender: address) (value: nat)
 pre-condition := not (paused = true)
  && get_big_map (Sender, ledger) = 
     Some (v: pair nat (map address nat))
  && v = Pair (n: nat) (m: map address nat) 
  && get_map (spender, m) = Some (k: nat) 
  && not (k = 0) && (value = 0)
 post-condition := get_big_map (Sender, post_ledger) = 
    Some (post_v:  (pair nat (map address nat))
  && post_v = Pair (post_n: nat) (post_m: map address nat) 
  && get_map (spender, post_m) = Some (post_k: nat) 
  && (post_k = 0) ;
\end{lstlisting}

@@

The Kolibri oracle maintains a state machine with two states: (1) \lstinline/IDLE/ and (2) \lstinline/WAITING_FOR_HARBINGER/. The \lstinline/getXtzUsdPrice/ entrypoint is permissible only when the oracle is in the \lstinline/IDLE/ state; invoking this entrypoint transitions the oracle to the \lstinline/WAITING_FOR_HARBINGER/ state. Conversely, the \lstinline/getXtzUsdPrice_callback/ entrypoint is allowable solely when the oracle is in the \lstinline/WAITING_FOR_HARBINGER/ state; invoking this entrypoint results in the oracle transitioning back to the \lstinline/IDLE/ state. The value of \lstinline/client/ is always set to \lstinline/none/ in the \lstinline/IDLE/ state and to \lstinline/some/ in the other state. We verify this property by examining the fail conditions of these two entrypoints and verifying the entrypoint relation between \lstinline/getXtzUsdPrice/ and \lstinline/getXtzUsdPrice_callback/.

The two subsequent properties dictate that if \lstinline/getXtzUsdPrice/ is successfully executed, the parameters must be stored in the \lstinline/client/ attribute. \\ If \lstinline/getXtzUsdPrice_callback/ is successful, there must be a transaction transferring the data to the client at the \lstinline/client/. These properties are checked by using pre- and post-conditions. Futhermore, the critical access control property is that the entrypoints \lstinline/setGovernorContract/ and \lstinline/setMaxDelaySec/ must only be invocable by the governor. Any attempts by unauthorized callers to these entrypoints should result in failure.

@@

We then add the predicate \lstinline/post_maxDelay <= 1800/ in the post condition of the entrypoint \lstinline/setMaxDataDelaySec/ as shown in Listing \ref{lst:setMaxDataDelaySec}, but then the pre and post condition test fails. This means there is no constraint that the data is less than 30 minutes. This constraint is possibly set up by the governor and specified in the \lstinline/MaxDataDelaySec/ field. However, users calling this contract should be aware that it is possible for the governor to set any value ($>=0$) as desired. This could lead to providing out-of-date rates.

\begin{lstlisting}[float=tp,captionpos=b,caption={The specification of the setMaxDataDelaySec entrypoint},label={lst:setMaxDataDelaySec},numbers=left]
entrypoint %setMaxDataDelaySec
 code := {...}
 parameter := delay: nat
 pre-condition := Sender = governor            
 post-condition := post_maxDelay = delay 
 && post_harbinger = harbinger && post_maxDelay <= 1800
\end{lstlisting}

The complete specifications and experimental results for the examples and case studies can be found in the appendix~\ref{sec:concl-sect-append} or accessed online at \cite{}.
\section{Symbolic Execution Model}
\label{sec:symbolic-execution-model} 
We address the changes required for symbolic execution of Michelson programs, focusing on a selection of instructions. The
appendix~\ref{sec:concl-sect-append} specifies further instructions,
but may be skipped by the reviewers.

\subsection{System Model}
\label{sec:system-model}
Our symbolic model of a program state consists of a stack where each element is a symbolic value, represented as a pair comprising a Michelson term and its type: \STACK\ = (\TermOne, \TYF) \STACKCONCAT\ (\TermTwo, \TYS) \STACKCONCAT\ \DOT\ \STACKCONCAT\ \EMPTYSTACK. Here, \ensuremath{[\ ]} represents the empty stack, and \ensuremath{hd :: tl} denotes a stack with \ensuremath{hd} as the top element and \ensuremath{tl} as the remaining stack. Let \INSTRUCTION\ = \InstructionOne; \InstructionTwo; \DOT\ \InstructionN\ be a sequence of instructions, and \PREDICATE\ be a path predicate expressed in conjunctive form, capturing the branching conditions.
\begin{definition}
A symbolic execution state is represented by a tuple \STATE\ =
[\INSTRUCTION, \STACK, \PREDICATE].
A system configuration, \SYSTEM\ = \{\STATEONE, \STATETWO,
\DOT, \STATEN \}, is a set of  states.
\end{definition}

>>>>

%
%\authorrunning{Ha Doan, P. Thiemann}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
%\institute{University of Freiburg, Germany \\
%  \email{\{doanha,thiemann\}@informatik.uni-freiburg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
We developed SCV, a verification tool geared towards contracts implemented in Michelson, the smart contract language of the Tezos blockchain. SCV utilizes symbolic execution to derive verification conditions that can be checked automatically against user specifications using an SMT solver. These specifications are written in a domain-specific language. We present the fundamental principles behind the design and development of SCV. By pruning states during symbolic execution, SCV effectively addresses the problem of state explosion. This approach has enabled us to conduct two case studies on real-life contracts, demonstrating the expressiveness of our tool.
\keywords{Smart Contract, Blockchain, Formal Verification, Symbolic Execution}
\end{abstract}
\section{Introduction}
\label{sec:introduction}
Since smart contracts are permanently stored on the blockchain, once deployed, a smart contract cannot be changed. Regrettably, numerous instances have been documented where vulnerabilities in smart contracts were identified by the blockchain community, resulting in substantial financial losses~\cite{dao,wallethack}. The distinctive execution model of smart contract languages has given rise to unforeseen errors due to a lack of familiarity with their complexities. These factors, combined with the immutability of contracts, underscore the need to ensure their correctness prior to deployment.

We have developed SCV, a formal verification tool for smart contracts on the Tezos blockchain \cite{tezos-whitepaper}, written in Michelson \cite{michelson}. SCV relies on symbolic execution backed by an SMT solver to simulate the implementation of the smart contract language, helping to detect subtle errors. Additionally, SCV includes a domain-specific language for precisely specifying smart contract properties, tailored to Michelson's execution model and inspired by Z3. Each specification consists of Michelson code, a precondition, and a postcondition with the usual partial correctness semantics. Our symbolic interpreter calculates a postcondition predicate from the specified precondition. The SMT solver Z3 \cite{z3} then checks that this predicate implies the specified postcondition.

While there is already a considerable body of work applying symbolic execution to the verification of smart contracts \cite{securify,manticore,kevm,park} (see Section~\ref{sec:related-work} for details), most of these efforts are designed for Ethereum, where standard techniques for symbolic execution of imperative programs are applicable. Such groundwork does not exist for Michelson, which is a functional language running on a stack machine (see Section~\ref{sec:background}) with a sophisticated instruction set that includes high-level loops. For that reason, we had to develop symbolic execution for Michelson from the ground up.

@@

Since Michelson operates on a stack, we assume that the input stack and the output stack each have only one element, as specified in the input term and the output term, respectively. The syntax allows us to specify both incomplete and complete smart contracts. For a complete smart contract, the output should be a pair consisting of an operation list and the storage's state after the completion of the code. For verification purposes, we elide the operation list and represent the output solely by the term of the value in the storage after execution concludes. %Michelson code, whether composed by users, obtained from the Tezos blockchain, or compiled from higher-level languages, can be seamlessly used within our framework.

@@

pre and postconditions, where the precondition states the assumption

@@

Let us have a look at the specification for the auction
contract that models an online auction in Listing \ref{lst:auction-contract-specification1}. The contract has two entry points, \lstinline|%bid| and \lstinline|%close|, which serve as functions for bidding and closing the contract, respectively. The smart contract storage is a pair containing, as its first element, a boolean value \lstinline|open| that indicates whether the contract is still open for bidding or already closed. The second element is another nested pair, where the first element \lstinline|highest| stores the highest bidder's address, and the second element \lstinline|owner| is the contract owner's address. The current highest bid is reflected in the balance of the contract before the next bid is invoked. We omit the Michelson code of the smart contract due to its length.  For the \lstinline/%bid/ entrypoint, the parameter is \lstinline/Unit/ of type \lstinline/unit/.
The current storage is described by the \lstinline|storage| phrase.

@@

\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification},label={lst:auction-contract-specification1},numbers=left]
mcontract auction = spec 
 storage := Pair (open: bool) 
                 (Pair (highest: address) (owner: address))
 entrypoint %bid
  code := {...}
  parameter := Unit 
  pre-condition := open = true && Amount > pre_Balance
   && get_contract (unit, highest) = Some (c: contract unit)
  post-condition := open = true && post_highest = Sender 
   && post_Balance = Amount && post_owner = owner 
   && transfer_token (Unit, pre_Balance, c);
 entrypoint %close
  ...
 (%create -> %bid) with ...
 | (%create  -> %close) with ...
 | (%bid -> %bid) with (Amount > pre_Balance) 
  && get_contract (unit, highest) = Some (c: contract unit)
 | (%bid -> %close) with (Sender = owner) 
  && get_contract (unit, owner) = Some (c: contract unit)
 | not (%close -> %bid) | not (%close -> %close)
\end{lstlisting}

%There are several properties that we want to guarantee for this contract. 
The first property is that if a user calls the \lstinline|%bid| entrypoint when the contract is still open for bidding and the bid is higher than the current bid (as specified in the \lstinline|pre-condition| field at line 7), and the address of the highest bidder can be cast to a real contract represented by the variable \lstinline|c| of type \lstinline|contract unit| on the blockchain (line 8), then the bid should succeed. In this case, the highest bidder must be updated to the sender, which is the address of the caller (line 9 in the \lstinline|post-condition| field), and the balance of the contract should be the amount sent along with the call (line 10). Moreover, the previous highest bidder should receive their bid back, which means the auction contract should transfer the \lstinline|pre_Balance| to the previous bidder, \lstinline|highest| (line 11). Another property guarantees that the auction owner's address should not be changed after calling the \lstinline/bid/ entrypoint (line 10) and the auction should be still open (line 9). 
\subsection{Entrypoint relations}
\label{sec:entrypoint-relations}
Since the order of smart contract
calls can be crucial, we specify ordering constraints among calls with
 entrypoint relations. An entrypoint relation can be specified as
\lstinline/%a -> %b with p/, which starts with the name
\lstinline/%a/ of the entrypoint and is followed by the keyword
\lstinline/with/ in case conditions are specified as predicates. The
following symbol \lstinline/->/ indicates that from the entrypoint
that satisfies the condition \lstinline/p/, it is possible  to call
the entrypoint \lstinline/%b/. One specification can
contain several entrypoint relations concatenated by
\lstinline/|/. There is a special entrypoint named \lstinline/create/
that indicates the contract state after deployment.

Let us return to the auction contract. It is important to guarantee
that  the bid call can happen successfully after another if the new bid is higher than the previous
one (lines 16-17). The  \lstinline/%close/ entrypoint can be called
anytime after bidding if the caller is the contract owner (lines 18-19). However, it
should not be possible to call the  \lstinline/%bid/ entrypoint after
the auction is closed, and, of course, it cannot be closed
again if it is already closed (line 20).  Furthermore, contract deployment sets the  \lstinline/open/ flag to
\TRUE, indicating that the contract is open for bidding, i.e., users
should be able to bid on the contract or the contract owner should be
able to close it. Otherwise, the owner's original
deposit is lost. This sequence  illustrates
the life cycle of the auction contract, which is depicted in Figure
\ref{fig:auction-life-cycle}. Other properties related to the \lstinline|%close| entrypoint and the life cycle are omitted here due to page limits. %, but the full specification can be found in the Appendix and is available online at \cite{scv}.

@@

\section{Static Checker}
\label{sec:static-checker-we}
The architecture of the static checker is illustrated in Figure~\ref{fig:architecture-of-static-checker}. Users write specifications in the SCV language, which are then parsed into an Abstract Syntax Tree (AST) in OCaml. The input term along with the code is passed to the symbolic interpreter, while the output term and the pre and postconditions are fed to the static checker. The symbolic interpreter, which is fully typed and rejects ill-typed programs, then runs symbolic execution on the code with the provided input, where the initial stack is derived from the input term.

@@

Consider the specification of the \lstinline/sub/ contract, the results of
the symbolic execution are either a failwith state with the path
condition \lstinline/x/ $<$ \lstinline/y/,  or a state in which the
final stack has only one element, which is a symbolic variable
\lstinline/sbvar/ of type integer, with the path condition stating
that the variable \lstinline/x/ is greater than or equal to
\lstinline/y/ and  \lstinline/sbvar/ is equal to
the substract of \lstinline/x/ and \lstinline/y/.

@@

\subsection{Output check}
\label{sec:output-check}
For an incomplete smart contract, the output term of the symbolic execution has the form of a term \Term\ and its type \TY. For a complete smart contract, it takes the form \PAIR\ \VOPERATIONLIST\ \VSTORAGE, \TPAIR\ (\TOPERATIONLIST) \TYS, where \VSTORAGE\ represents the storage and \VOPERATIONLIST\ represents the operation list. Given an output term specified as $\tau$, the checker scans all reachable final states that do not contain a failwith stack and checks whether the output term matches. In the former case, it matches $\tau$ with the term \Term. In the latter case, it matches with \VSTORAGE.

For instance, in the \lstinline|sub| contract, the output check ignores the states where the failwith occurs and verifies whether the output term \lstinline|z| matches with the symbolic output term \lstinline/sbvar/ in the other state. It succeeds for this example as expected. However, this test could fail if, for example, the output term is incorrectly specified as a variable of a type that does not match the actual result.

\subsection{Fail conditions}
\label{sec:fail-conditions}
A Michelson contract can fail due to conditions introduced by the programmer, triggered by the \FAILWITH\ instruction along with an error message. Failures play a crucial role in smart contract programming to ensure that certain conditions lead to contract termination, such as unauthorized caller attempts (access control properties) or invalid arguments. Various smart contract standards, including \cite{erc,fa}, mandate specific fail conditions.

@@

Any contract aiming to implement the FA 1.2 standard must incorporate the functions specified in Listing~\ref{lst:fa12-interface}. We verify USDtz \cite{usdtzimplemetation,tzstatsusdtz}, an implementation of the FA 1.2 standard. The goal is to ensure that the implementation aligns with the standard, verifying its correctness. USDtz is designed to maintain a stable value by pegging it to a fiat currency like USD or EUR, and it has been operational on the Tezos blockchain for a significant period.

@@

The \lstinline/transfer/ entrypoint credits the account associated with the address provided in the \lstinline/to/ parameter while debiting the account corresponding to the \lstinline/from/ parameter with the specified \lstinline/value/ amount. According to the FA 1.2 standard, this call must fail under two specific conditions: (1) The entrypoint must fail with the error \lstinline/NotEnoughBalance/ if the sender's account lacks sufficient funds. (2) If the sender's account does not have permission to withdraw the specified \lstinline/value/ amount, the call will fail with the \lstinline/NotEnoughAllowance/ error. To verify these requirements, we utilize the failwith condition feature of our tool and examine the returned results, one of which is outlined in Listing~\ref{lst:transfer-udstz-contract}. The fail condition results confirm these stipulations.
\begin{lstlisting}[float,captionpos=b,caption={The \lstinline/NotEnoughBalance/ fail condition for the \lstinline/transfer/ entrypoint},label={lst:transfer-udstz-contract},numbers=left]
not (paused = true) 
&& get_big_map (from, ledger) = Some sbvar 
&& sbvar = Pair sbvar1 sbvar2 
&& not (sbvar1 >= value) : 'NotEnoughBalance'
\end{lstlisting}
Another essential property is ensuring that the allowance of the sender's account is decremented and the receiver's account is incremented by the \lstinline/value/ that has been sent. This specification is verified to ensure compliance with the contract's intended behavior.

The \texttt{approve} entrypoint grants the spender the right to transfer a specified amount from the owner's account. The contract prohibits changing the allowance value from a non-zero value to another non-zero value to prevent a known attack vector \cite{attack-vector}. This attack exploits the ability to spend the entire allowance before an update and then again after, resulting in a larger amount being spent than allowed. Therefore, the contract call must fail if both the parameter \lstinline/value/ and the current allowance of \lstinline/spender/ are non-zero. We verify this critical property by generating fail conditions to check whenever such a non-zero value update results in a fail condition as Listing \ref{lst:approve-udstz-contract-z}.
\begin{lstlisting}[float=tp,captionpos=b,caption={The non-zero value fail condition for the \lstinline/approve/ entrypoint},label={lst:approve-udstz-contract-z},numbers=left]
not (paused = true) 
&& get_big_map(Sender, ledger) = Some sbvar 
&& sbvar = Pair sbvar1 sbvar2 
&& get_map(spender, sbvar2) = Some sbvar3 
&& not(value = 0) && not(sbvar3 = 0): 'UnsafeAllowanceChange'
\end{lstlisting}

When the update is allowed, the updated value must be correctly recorded in the storage. This can be divided into two sub-properties: (1) if the parameter \lstinline/value/ is zero, then the corresponding allowance value of the \lstinline/spender/ from the owner (which is \SENDER\ in this call) should be zero, and (2) if \lstinline/value/ is non-zero and the allowance of the \lstinline/spender/ is zero, then the \lstinline/value/ should be recorded in the storage after the call. We specify these properties using pre- and post-conditions, as illustrated in Listing~\ref{lst:specification-approve-p1} for the first property.

@@

We specified and verified all the properties mentioned above and others. The results confirm that these properties are guaranteed, ensuring the correctness of the smart contract code.
\begin{lstlisting}[float=tp,captionpos=b,caption={Specification of the \lstinline/approve/ entrypoint (Property 1)},label={lst:specification-approve-p1},numbers=left]
entrpoint %approve
 parameter := Pair (spender: address) (value: nat)
 pre-condition := not (paused = true)
  && get_big_map (Sender, ledger) = 
     Some (v: pair nat (map address nat))
  && v = Pair (n: nat) (m: map address nat) 
  && get_map (spender, m) = Some (k: nat) 
  && not (k = 0) && (value = 0)
 post-condition := get_big_map (Sender, post_ledger) = 
    Some (post_v:  (pair nat (map address nat))
  && post_v = Pair (post_n: nat) (post_m: map address nat) 
  && get_map (spender, post_m) = Some (post_k: nat) 
  && (post_k = 0) ;
\end{lstlisting}

@@

The Kolibri oracle maintains a state machine with two states: (1) \lstinline/IDLE/ and (2) \lstinline/WAITING_FOR_HARBINGER/. The \lstinline/getXtzUsdPrice/ entrypoint is permissible only when the oracle is in the \lstinline/IDLE/ state; invoking this entrypoint transitions the oracle to the \lstinline/WAITING_FOR_HARBINGER/ state. Conversely, the\\ \lstinline/getXtzUsdPrice_callback/ entrypoint is allowable solely when the oracle is in the \lstinline/WAITING_FOR_HARBINGER/ state; invoking this entrypoint results in the oracle transitioning back to the \lstinline/IDLE/ state. The value of \lstinline/client/ is always set to \lstinline/none/ in the \lstinline/IDLE/ state and to \lstinline/some/ in the other state. We verify this property by examining the fail conditions of these two entrypoints and verifying the entrypoint relations between \lstinline/getXtzUsdPrice/ and \lstinline/getXtzUsdPrice_callback/.

The two subsequent properties dictate that if \lstinline/getXtzUsdPrice/ is successfully executed, the parameters must be stored in the \lstinline/client/ attribute. \\ If \lstinline/getXtzUsdPrice_callback/ is successful, there must be a transaction transferring the data to the client at the \lstinline/client/. These properties are checked by using pre and postconditions. Futhermore, the critical access control property is that the entrypoints \lstinline/setGovernorContract/ and \lstinline/setMaxDelaySec/ must only be invocable by the governor. Any attempts by unauthorized callers to these entrypoints should result in failure.

@@

We then add the predicate \lstinline/post_maxDelay <= 1800/ in the post condition of the entrypoint \lstinline/setMaxDataDelaySec/ as shown in Listing \ref{lst:setMaxDataDelaySec1}, but then the pre and post condition test fails. This means there is no constraint that the data is less than 30 minutes. This constraint is possibly set up by the governor and specified in the \lstinline/MaxDataDelaySec/ field. However, users calling this contract should be aware that it is possible for the governor to set any value ($>=0$) as desired. This could lead to providing out-of-date rates.

\begin{lstlisting}[float=tp,captionpos=b,caption={The specification of the \lstinline/setMaxDataDelaySec/ entrypoint},label={lst:setMaxDataDelaySec1},numbers=left]
entrypoint %setMaxDataDelaySec
 code := {...}
 parameter := delay: nat
 pre-condition := Sender = governor            
 post-condition := post_harbinger = harbinger 
 && post_governor = gover
 && post_maxDelay = delay && post_maxDelay <= 1800
\end{lstlisting}

The complete specifications and experimental results for the examples and case studies can be found in the appendix~\ref{sec:appendix} or accessed online at \cite{scv}.
\section{Symbolic Execution Model}
\label{sec:symbolic-execution-model} 
We address the changes required for symbolic execution of Michelson programs, focusing on a selection of instructions. The
appendix~\ref{sec:concl-sect-append} specifies further instructions,
but may be skipped by the reviewers.
